\Large The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events

### 2023-09-23 paper/main.tex
finished a complete round of edits (main text + supplement)

\begin{abstract} {\footnotesize{ How much can we infer about the past and
future, given our knowledge of the present? Unlike temporally symmetric
inferences about simple sequences, inferences about our own lives are
asymmetric: we are better able to infer the past than the future, since we
remember our past but not our future (i.e., the psychological arrow of time).
What happens when both the past and future are unobserved, as when we make
inferences about \textit{other} people's lives? We had participants in two
experiments view segments of two character-driven television dramas. They wrote
out what would happen just before or after each just-watched segment.
Participants were better at inferring past (versus future) events. This
asymmetry was driven by participants’ reliance on characters’ conversational
references in the narrative, which tended to favor the past. We also carried
out a meta analysis to estimate the prevalence of these asymmetries in hundreds
of millions of dialogues from television shows, popular movies, novels, and
written and spoken natural conversations. We found that, on average, references
to the past are roughly 1.5--2 times more prevalent than references to the
future. Our work reveals a temporal asymmetry in how observations of other
people’s behaviors can inform us about the past and future.

@@

One key difference between simple artificial sequences and more naturalistic
(real or narrative) sequences is that naturalistic sequences often incorporate
other people. Despite the past and future being equally unknown to \textit{the
observer} prior to the current moment, other people, and realistic characters
in narratives, have their own psychological arrows of time. Specifically, they
have memories of their own pasts. Other people's asymmetric knowledge about
their \textit{own} pasts and futures might affect their behaviors (e.g.,
conversations). In turn, this might provide time-asymmetric clues that favor
the past~\citep[e.g., other people might talk more about their own pasts than
their futures; ][]{DemiEtal18}. If observers leverage these clues from other
people's asymmetric knowledge, then observers should also be better at
inferring the past (versus the future) of other people's lives. Alternatively,
inferences about other people's lives may be more like inferences about
artificial statistical sequences~\citep[e.g., perhaps solely relying on
statistical regularities like event schemas, scripts, or situation
models][]{RadvCope06, ZwaaRadv98, BoweEtal79, RangRitc12, BaldEtal18}. If so,
then the accuracy of inferences about the past and the future of others' lives
should be approximately equal. We note that the aforementioned authors make no
specific claims about temporal symmetries or asymmetries. Rather, we claim that
statistical regularities might \textit{imply} symmetry (e.g., if you are on
step $n$ of an unfolding schema, this suggests you have just completed step $n
- 1$ and that you are next likely to encounter step $n + 1$).

@@

For each retrodiction and prediction, participants were asked to generate at
least one, and not more than three, responses that constituted ``the sorts of
things [the participant would] expect to have remembered if [they] had watched
the [target] segment.'' They were asked to generate multiple responses only if
those additional responses were (in their judgement) of equal likelihood to
occur. On average, participants generated 1.08 responses per prompt; therefore
we chose to consider only participants' first (``most probable'' or ``most
important'') responses to each prompt. We also discarded a small number ($n =
20$) of character-cued responses that did not contain references to all cued
characters, along with one additional response due to the participant's
misunderstanding of the task instructions during that trial. We carried out our
analyses on the remaining 2084 retrodiction, prediction, and recall responses.

@@

  \caption{\textbf{Retrodiction, prediction, and recall performance by
  experimental condition.} \textbf{A. Methods schematic.} For each
  retrodiction, prediction, and recall response, we calculated the hit rate for
  events in the target segment, the response precision (see \textit{Methods}),
  and the response convergence across participants (see \textit{Methods}).
  \textbf{B. Target event hit rate.} Mean proportions of target events that
  were contained in participants' responses, for each response type, averaged
  across target segments. \textbf{C. Response precision.} Mean precisions of
  participants' responses, for each response type, averaged across target
  segments. The horizontal lines denote the mean pairwise semantic similarities
  (see \textit{Methods}) across recall responses (re(R): orange; re(P): blue).
  \textbf{D. Response convergence.} Mean (across-participant) convergence of
  participants' responses, for each response type, averaged across target
  segments. All panels: error bars denote bootstrapped 95\% confidence
  intervals. Asterisks indicate significance in the (generalized) linear mixed
  models: * denotes $p < 0.05$ and ** denotes $p < 0.01$. See
  Figure~\targetAsymmetries for analogous results from our replication
  experiment.}

@@

We used two general approaches to assess the quality of participants' responses
(see \textit{Methods}, Figs.~\ref{fig:result1}A). One approach entailed
manually annotating events in the video and counting the number of matched
events in participants' responses. We identified a total of 117 unique events
reflected across the 22 video segments (range: 3--9 per segment; see
\textit{Methods}, Table \stimDescription). We assigned one ``point'' to each of
these video events. We also identified 23 additional events in participants'
responses that were either summaries of several events or that were partial
matches to the manually identified video events. We assigned 0.5 point to each
of these additional events. This point system enabled us to compute the numbers
and proportions (\textit{hit rates}) of correctly retrodicted, predicted, and
recalled events contained in each response. Our second approach entailed using
a natural language processing model~\citep{CerEtal18} to embed annotations and
responses in a 512-dimensional feature space. This approach was designed to
capture conceptual overlap between responses that were not necessarily tied to
specific events. To quantify this conceptual overlap, we computed the
similarities between the embeddings of different sets of responses. Following
\cite{HeusEtal21}, we defined the \textit{precision} of each participants'
retrodictions or predictions about a target segment as the median cosine
similarities between the embeddings of (a) the participant's retrodiction or
prediction response for the target segment and (b) each \textit{other}
participant's recalls of the same segment. In other words, precision is
designed to measure the extent to which retrodictions and predictions captured
the conceptual content that (other) participants remembered. We also developed
a related measure, which we call \textit{convergence}, to characterize response
similarities across participants. In particular, we defined convergence as the
mean cosine similarity between the embeddings of a participant's responses to a
target segment and all other participants' responses (of the same type) to the
same segment. We analyzed the data using generalized linear mixed models, with
participant and stimulus (e.g., target segment) identities as crossed random
effects (see \textit{Methods}).

First we sought to validate a main effect of response type (i.e., uncued
responses, character-cued responses, and recalls), irrespective of the temporal
direction (retrodiction versus prediction). Across these three types of
responses, participants have access to increasing amounts of information about
the target segment. Therefore, across these response types, we hypothesized
that participants' responses should become both more accurate and more
convergent across individuals. Consistent with this hypothesis, participants'
character-cued retrodictions and predictions were associated with higher target
event hit rates than uncued retrodictions and predictions (odds ratio (OR):
2.65, $Z = 4.24$, $p < 0.001$, 95\% confidence interval (CI): 1.69 to 4.16;
Fig.~\ref{fig:result1}B). These character-cued responses were also more precise
($b = 0.13$, $t(18.1) = 9.43$, $p < 0.001$, CI: 0.10 to 0.16;
Fig.~\ref{fig:result1}C) and convergent across individuals ($b = 0.11$,
$t(18.6) = 6.21$, $p < 0.001$, CI: 0.07 to 0.15; Fig.~\ref{fig:result1}D).
Relative to character-cued responses, participants' recalls showed higher
target event hit rates (OR = 21.83, $Z = 10.61$, $p < 0.001$, CI: 12.35 to
38.59) and were more convergence across individuals ($b = 0.20$, $t(19.4) =
9.10$, $p < 0.001$, CI: 0.16 to 0.25). These results are consistent with the
common-sense notion that access to more information about a target segment
yields better performance (i.e., higher hit rates, precision, and convergence
across individuals). These findings also held for our replication experiment
(Fig.~\targetAsymmetries; hit rates of character-cued vs. uncued responses: OR:
XXX, $Z = XXX$, $p = XXX$, 95\% confidence interval (CI): XXX to XXX;
precisions of character-cued vs. uncued responses: $b = XXX$, $t(XXX) = XXX$,
$p = XXX$, CI: XXX to XXX; convergence of character-cued vs. uncued responses:
$b = XXX$, $t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX).

@@

  \caption{\textbf{Retrodictions and predictions of temporally near and distant
  events.} \textbf{A. Illustration of annotation approach.} For each uncued
  retrodiction and prediction response in our main experiment, we calculated
  the number of (retrodicted or predicted) events as a function of temporal
  distance from the target segment, or \textit{lag}. Onscreen (explicit) events
  are tagged using integer-valued lags, whereas offscreen (implicit) events are
  tagged using half-step lags ($\pm 0.5$, $\pm 1.5$, etc.). \textbf{B. Number
  of events hit in participants' uncued retrodictions and predictions for each
  event type.} Here we separated events we identified in participants'
  responses according to whether they occurred in the target segment (lags of
  $\pm 1$), during the interval between the target segment and the just-watched
  segment (lags of $\pm 0.5$), at longer temporal distances ($|\mathrm{lag}| >
  1$), or were incorrect (unmatched with any past or future events in the
  narrative). The counts displayed in the panel are averaged across
  just-watched segments. \textbf{C. Number of events hit as a function of
  temporal distance.} Here the (across-segment) mean numbers of events hit in
  participants' uncued retrodictions (orange) and predictions (blue) are
  displayed as a function of temporal distance to the just-watched segment
  (lag). Error bars denote bootstrapped 95\% confidence intervals. Colors
  denote temporal direction (orange: past; blue: future) and distance (darker
  shading: onscreen events from segments adjacent to the target segment;
  lighter shading: offscreen events). See Figure~\hitRates for an analogous
  presentation of results from our replication study.}

@@

The above analyses were focused solely on the target segment (i.e.,
retrodiction of segment $n$ after watching segments $(n+1)... 11$, or
prediction of segment $n$ after watching segments $1 ...(n-1)$). We wondered
whether participants' responses might also contain longer-range information
about preceding or proceeding events. In order to carry out this analysis
properly, we reasoned that participants might reference past or future events
that were \textit{implied} to have occurred offscreen, but not explicitly shown
onscreen. For example, a character in location A during one scene might appear
in location B during the immediately following scene. Although it wasn't shown
onscreen, we can infer that the character traveled between locations A and B
sometime between the time intervals separating the scenes~\citep{Bord08}. In
all, we manually identified a set of 74 \textit{implicit} offscreen events that
were implied to have occurred given what was (explicitly) depicted onscreen
(Fig.~\ref{fig:result2}A), plus one additional partial event and one additional
summary event. We defined the just-watched segment as having a \textit{lag} of
0. We assigned the target segment of a participant's retrodiction or prediction
(i.e., the immediately preceding or proceeding segment) a lag of -1 or +1,
respectively. The segment following the next was assigned a lag of 2, and so
on. We tagged offscreen events using half steps. For example, an offscreen
event that occurred after the prior segment but before the just-watched segment
would be assigned a lag of -0.5.

@@

What might be driving participants to retrodict further and more accurately
into the unobserved past, compared with their predictions of the unobserved
future? By inspecting the video content, we noticed that characters in the
television show frequently referenced both past events and (planned or
predicted) future events in their spoken conversations. We wondered whether the
characters' references might show temporal asymmetries that might explain
participants' behaviors. Across all of the characters' conversations, and
across all of the video segments, we manually identified a total of 82
references to past or future events (i.e., that occurred onscreen or offscreen
before or after the events depicted in the current segment;
Figs.~\ref{fig:result3}A, \references A, \characterRefs). Characters in our
main experiment's stimulus tended to reference the past (52 references) more
than the future (30 references), consistent with previous
work~\citep{DemiEtal18}. References to the past were also skewed to more
temporally distant events compared with references to the future
(Figs.~\ref{fig:result3}B,~\references B,~\characterRefs). These asymmetries
also held for characters in the replication experiment's stimulus
(Fig.~\ref{fig:meta-analysis}). These observations indicate that the characters
in the stimulus display a preference for the past (versus future) in their
conversations. Might this asymmetry be driving the asymmetries in participants'
retrodictions versus predictions?

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.9\textwidth]{results3}

  \caption{\textbf{Characters' references drive participants' retrodiction and
  prediction performance.} \textbf{A. Illustration of annotation approach.} We
  manually annotated references to events in past or future segments in
  characters' spoken conversations. We matched each such reference with its
  corresponding storyline event (and its corresponding segment number for
  onscreen events, or half-step segment number for offscreen events). We then
  tracked the hit rate separately for referenced versus unreferenced events in
  participants' uncued retrodictions and predictions. \textbf{B. Reference rate
  as a function of lag.} Across all possible just-watched segments (lag 0), the
  bar heights denote the average proportions of events referenced in other past
  (orange, negative lags) or future (blue, positive lags) segments in our main
  experiment's stimulus. \textbf{C. Difference in hit rates between all events
  and unreferenced events.} To highlight the effect of characters' references
  to past and future events on participants' retrodictions and predictions,
  here we display the difference in across-segment mean hit rates between all
  events and unreferenced events, as a function of temporal distance (lag) to
  the just-watched segment. \textbf{D. Hit rates for unreferenced events.} The
  average response hit rates for unreferenced events are displayed as a
  function of temporal distance to the just-watched segment. Error bars denote
  bootstrapped 95\% confidence intervals. Panels B--D: colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{E. Hit rates and counts of
  referenced and unreferenced events.} As a function of temporal distance to
  the just-watched segment, the sub-panels display the across-segment mean
  numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red) and
  unreferenced (gray) events that participants hit (darker shading) or missed
  (lighter shading) in their uncued retrodictions (top sub-panel) and uncued
  predictions (bottom sub-panel). Intuitively, the widths of the rectangles at
  each lag denote the total number of events at each possible lag. The darker
  shading denotes the proportions of events that participants retrodicted or
  predicted, and the lighter shading denotes the proportions of events that
  participants ``missed'' in their responses. For an analogous presentation of
  results from the replication experiment, see Fig.~\characterRefs.}

@@

  \caption{\textbf{Referenced events are associated with higher hit rates, but
  referring events are not.} \textbf{A. Illustration of annotation approach.}
  We extended the annotation procedure depicted in Figure~\ref{fig:result3}A to
  also label which events in our main experiment's stimuli \textit{contained}
  references to events in other segments. \textbf{B. Referenced versus
  referring events.} During event $i$, when a character makes a reference to
  another event ($j$), we define $i$ as the \textit{referring} event and $j$ as
  the \textit{referenced} event. \textbf{C. Referring rate as a function of
  lag.} Across all possible just-watched segments (lag 0), the bar heights
  denote the across-segment mean proportions of events containing references to
  events in other past (orange, negative lags) or future (blue, positive lags)
  segments. The bar colors are described in the Figure~\ref{fig:result2}
  caption. \textbf{D. Hit rates and counts of referenced, referring, and other
  events.} As a function of temporal distance to the just-watched segment, the
  sub-panels display the numbers ($x$-axes) and hit rates ($y$-axes) of
  referenced (red), referring (green), and other (gray) events that
  participants hit (darker shading) or missed (lighter shading) in their uncued
  retrodictions (top sub-panel) and uncued predictions (bottom sub-panel). For
  a display of analogous results from our replication experiment see
  Figure~\referringReferenced.}

@@

The preceding analyses show that when characters reference past or future
events, those referenced events, and other events that are temporally adjacent
to the referenced events, are more likely to be retrodicted and predicted. In
other words, referring to a past or future event in conversation leads to a
``boost'' in that event's hit rate. We wondered whether this boost was
bi-directional. In particular: when a character refers (during a
\textit{referring event}) to another event (i.e., the \textit{referenced
event}), does this boost only the referenced event's hit rate, or does the
referring event also receive a boost? We labeled each event as a ``referring
event,'' a ``referenced event,'' or a ``other event'' (i.e., not referring or
referenced; Fig.~\ref{fig:result5}A, B). We limited our analysis to references
to onscreen (explicit) events. Consistent with our analysis of the proportions
of referenced events (Fig.~\ref{fig:result3}B), the proportions of
\textit{referring} events exhibited a \textit{forward} temporal asymmetry
(Fig.~\ref{fig:result5}C). Controlling for absolute lag, we found that
referring events were associated with lower hit rates than referenced events in
our main experiment (uncued retrodiction: OR = 0.03, $Z = -4.81$, $p < 0.001$,
CI: 0.01 to 0.11; uncued prediction: OR = 0.04, $Z = -5.84$, $p < 0.001$, CI:
0.01 to 0.12; Fig.~\ref{fig:result5}D) and had no reliable differences in hit
rates compared with other events (uncued retrodiction: OR = 0.37, $Z = -1.46$,
$p = 0.15$, CI: 0.10 to 1.41; uncued prediction: OR = 2.16, $Z = 1.68$, $p =
0.09$, CI: 0.88 to 5.30). We also observed this phenomenon in our replication
experiment (referenced events, uncued retrodiction: OR = XXX, $Z = XXX$, $p =
XXX$, CI: XXX to XXX; referenced events, uncued prediction: OR = XXX, $Z =
XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued retrodiction: OR = XXX,
$Z = XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued prediction: OR =
XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX;Fig.~\referringReferenced). Taken
together, this indicates that only referenced events received a hit rate boost
(relative to other events), suggesting that the retrodictive and predictive
benefits of references are directed (i.e., asymmetric).

@@

An additional ambiguous case arose in one participant's responses pertaining to
segment 12, storyline 2, whereby the participant correctly identified an
onscreen event that had not been included in our original annotations. To
account for this participant's response, we retroactively added that event to
our annotations of that segment. We also identified and counted unmatched
events in participants' responses (i.e., events that did not match any
annotated events). Cases where the two coders' independent scoring disagreed
were resolved through discussions between the two coders.

@@

In principle, a given event could receive multiple labels. For example, during
event $A$, a character might speak about another event, $B$, during which a
reference to a third event ($C$) was made. In this scenario, event $B$ could be
both a ``referring event'' ($B \rightarrow C$) \textit{and} a referenced event
($A \rightarrow B$). In practice, however, this scenario was quite rare,
accounting for only one out of a total of 30 onscreen events.

>>>>

\begin{abstract} {\footnotesize{ How much can we infer about the past and
future, given our knowledge of the present? Unlike temporally symmetric
inferences about simple sequences, inferences about our own lives are
asymmetric: we are better able to infer the past than the future, since we
remember our past but not our future (i.e., the psychological arrow of time).
What happens when both the past and future are unobserved, as when we make
inferences about \textit{other} people's lives? We had participants in two
experiments view segments of two character-driven television dramas. They wrote
out what would happen just before or after each just-watched segment.
Participants were better at inferring past (versus future) events. This
asymmetry was driven by participants’ reliance on characters’ conversational
references in the narrative, which tended to favor the past. We also carried
out a meta analysis to estimate the prevalence of these asymmetries in hundreds
of millions of dialogues from television shows, popular movies, novels, and
written and spoken natural conversations. We found that, on average, references
to the past are roughly 1.5--2 times more prevalent in human conversations than
references to the future. Our work reveals a temporal asymmetry in how
observations of other people’s behaviors can inform us about the past and
future.

@@

One key difference between simple artificial sequences and more naturalistic
(real or narrative) sequences is that naturalistic sequences often incorporate
other people. Despite the past and future being equally unknown to \textit{the
observer} prior to the current moment, other people, and realistic characters
in narratives, have their own psychological arrows of time. Specifically, they
have memories of their own pasts. Other people's asymmetric knowledge about
their \textit{own} pasts and futures might affect their behaviors (e.g.,
conversations). In turn, this might provide time-asymmetric clues that favor
the past~\citep[e.g., other people might talk more about their own pasts than
their futures; ][]{DemiEtal18}. If observers leverage these clues from other
people's asymmetric knowledge, then observers should also be better at
inferring the past (versus the future) of other people's lives. Alternatively,
inferences about other people's lives may be more like inferences about
artificial statistical sequences~\citep[e.g., perhaps solely relying on
statistical regularities like event schemas, scripts, or situation
models;][]{RadvCope06, ZwaaRadv98, BoweEtal79, RangRitc12, BaldEtal18}. If so,
then the accuracy of inferences about the past and the future of others' lives
should be approximately equal. We note that the aforementioned authors make no
specific claims about temporal symmetries or asymmetries. Rather, we claim that
statistical regularities might \textit{imply} symmetry (e.g., if you are on
step $n$ of an unfolding schema, this suggests you have just completed step $n
- 1$ and that you are likely to next encounter step $n + 1$).

@@

For each retrodiction and prediction, participants were asked to generate at
least one, and not more than three, responses that constituted ``the sorts of
things [the participant would] expect to have remembered if [they] had watched
the [target] segment.'' They were asked to generate multiple responses only if
those additional responses were (in their judgement) of equal likelihood to
occur. On average, participants in our main experiment generated 1.08 responses
per prompt; therefore we chose to consider only participants' first (``most
probable'' or ``most important'') responses to each prompt. We also discarded a
small number ($n = 20$) of character-cued responses that did not contain
references to all cued characters, along with one additional response due to
the participant's misunderstanding of the task instructions during that trial.
We carried out our analyses on the remaining 2084 retrodiction, prediction, and
recall responses. (Our replication experiment analyses were carried out on XXX
responses.)

@@

  \caption{\textbf{Retrodiction, prediction, and recall performance by
  experimental condition in our main experiment.} \textbf{A. Methods
  schematic.} For each retrodiction, prediction, and recall response, we
  calculated the hit rate for events in the target segment, the response
  precision (see \textit{Methods}), and the response convergence across
  participants (see \textit{Methods}). \textbf{B. Target event hit rate.} Mean
  proportions of target events that were contained in participants' responses,
  for each response type, averaged across target segments. \textbf{C. Response
  precision.} Mean precisions of participants' responses, for each response
  type, averaged across target segments. The horizontal lines denote the mean
  pairwise semantic similarities (see \textit{Methods}) across recall responses
  (re(R): orange; re(P): blue). \textbf{D. Response convergence.} Mean
  (across-participant) convergence of participants' responses, for each
  response type, averaged across target segments. All panels: error bars denote
  bootstrapped 95\% confidence intervals. Asterisks indicate significance in
  the (generalized) linear mixed models: * denotes $p < 0.05$ and ** denotes $p
  < 0.01$. See Figure~\targetAsymmetries~for analogous results from our
  replication experiment.}

@@

We used two general approaches to assess the quality of participants' responses
(see \textit{Methods}, Figs.~\ref{fig:result1}A). One approach entailed
manually annotating events in the video and counting the number of matched
events in participants' responses. We identified a total of 117 unique events
reflected across the 22 video segments in our main experiment (range: 3--9 per
segment; see \textit{Methods}, Table \stimDescription). We assigned one
``point'' to each of these video events. We also identified 23 additional
events in participants' responses that were either summaries of several events
or that were partial matches to the manually identified video events. We
assigned 0.5 point to each of these additional events. This point system
enabled us to compute the numbers and proportions (\textit{hit rates}) of
correctly retrodicted, predicted, and recalled events contained in each
response. Our second approach entailed using a natural language processing
model~\citep{CerEtal18} to embed annotations and responses in a 512-dimensional
feature space. This approach was designed to capture conceptual overlap between
responses that were not necessarily tied to specific events. To quantify this
conceptual overlap, we computed the similarities between the embeddings of
different sets of responses. Following \cite{HeusEtal21}, we defined the
\textit{precision} of each participants' retrodictions or predictions about a
target segment as the median cosine similarities between the embeddings of (a)
the participant's retrodiction or prediction response for the target segment
and (b) each \textit{other} participant's recalls of the same segment. In other
words, precision is designed to measure the extent to which retrodictions and
predictions captured the conceptual content that (other) participants
remembered. We also developed a related measure, which we call
\textit{convergence}, to characterize response similarities across
participants. In particular, we defined convergence as the mean cosine
similarity between the embeddings of a participant's responses to a target
segment and all other participants' responses (of the same type) to the same
segment. We analyzed the data using generalized linear mixed models, with
participant and stimulus (e.g., target segment) identities as crossed random
effects (see \textit{Methods}).

First we sought to validate a main effect of response type (i.e., uncued
responses, character-cued responses, and recalls), irrespective of the temporal
direction (retrodiction versus prediction). Across these three types of
responses, participants have access to increasing amounts of information about
the target segment. Therefore, across these response types, we hypothesized
that participants' responses should become both more accurate and more
convergent across individuals. Consistent with this hypothesis, participants'
character-cued retrodictions and predictions were associated with higher target
event hit rates than uncued retrodictions and predictions in our main
experiment (odds ratio (OR): 2.65, $Z = 4.24$, $p < 0.001$, 95\% confidence
interval (CI): 1.69 to 4.16; Fig.~\ref{fig:result1}B). These character-cued
responses were also more precise ($b = 0.13$, $t(18.1) = 9.43$, $p < 0.001$,
CI: 0.10 to 0.16; Fig.~\ref{fig:result1}C) and convergent across individuals
($b = 0.11$, $t(18.6) = 6.21$, $p < 0.001$, CI: 0.07 to 0.15;
Fig.~\ref{fig:result1}D). Relative to character-cued responses, participants'
recalls showed higher target event hit rates (OR = 21.83, $Z = 10.61$, $p <
0.001$, CI: 12.35 to 38.59) and were more convergence across individuals ($b =
0.20$, $t(19.4) = 9.10$, $p < 0.001$, CI: 0.16 to 0.25). These results are
consistent with the common-sense notion that access to more information about a
target segment yields better performance (i.e., higher hit rates, precision,
and convergence across individuals). These findings also held for our
replication experiment (Fig.~\targetAsymmetries; hit rates of character-cued
vs. uncued responses: OR: XXX, $Z = XXX$, $p = XXX$, 95\% confidence interval
(CI): XXX to XXX; precisions of character-cued vs. uncued responses: $b = XXX$,
$t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX; convergence of character-cued vs.
uncued responses: $b = XXX$, $t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX).

@@

  \caption{\textbf{Retrodictions and predictions of temporally near and distant
  events.} \textbf{A. Illustration of annotation approach.} For each uncued
  retrodiction and prediction response in our main experiment, we calculated
  the number of (retrodicted or predicted) events as a function of temporal
  distance from the target segment, or \textit{lag}. Onscreen (explicit) events
  are tagged using integer-valued lags, whereas offscreen (implicit) events are
  tagged using half-step lags ($\pm 0.5$, $\pm 1.5$, etc.). \textbf{B. Number
  of events hit in participants' uncued retrodictions and predictions for each
  event type.} Here we separated events we identified in participants'
  responses according to whether they occurred in the target segment (lags of
  $\pm 1$), during the interval between the target segment and the just-watched
  segment (lags of $\pm 0.5$), at longer temporal distances ($|\mathrm{lag}| >
  1$), or were incorrect (unmatched with any past or future events in the
  narrative). The counts displayed in the panel are averaged across
  just-watched segments. \textbf{C. Number of events hit as a function of
  temporal distance.} Here the (across-segment) mean numbers of events hit in
  participants' uncued retrodictions (orange) and predictions (blue) are
  displayed as a function of temporal distance to the just-watched segment
  (lag). Error bars denote bootstrapped 95\% confidence intervals. Colors
  denote temporal direction (orange: past; blue: future) and distance (darker
  shading: onscreen events from segments adjacent to the target segment;
  lighter shading: offscreen events). See Figure~\hitRates~for an analogous
  presentation of results from our replication study.}

@@

The above analyses were focused solely on the target segment (i.e.,
retrodiction of segment $n$ after watching segments $(n+1)... 11$, or
prediction of segment $n$ after watching segments $1 ...(n-1)$). We wondered
whether participants' responses might also contain longer-range information
about preceding or proceeding events. In order to carry out this analysis
properly, we reasoned that participants might reference past or future events
that were \textit{implied} to have occurred offscreen, but not explicitly shown
onscreen. For example, a character in location A during one scene might appear
in location B during the immediately following scene. Although it wasn't shown
onscreen, we can infer that the character traveled between locations A and B
sometime between the time intervals separating the scenes~\citep{Bord08}. In
all, we manually identified a set of 74 \textit{implicit} offscreen events in
our main experiment's stimuli that were implied to have occurred given what was
(explicitly) depicted onscreen (Fig.~\ref{fig:result2}A), plus one additional
partial event and one additional summary event. We applied the same procedure
to our replication experiment's stimuli and identified XXX implicit offscreen
events. We defined the just-watched segment as having a \textit{lag} of 0. We
assigned the target segment of a participant's retrodiction or prediction
(i.e., the immediately preceding or proceeding segment) a lag of -1 or +1,
respectively. The segment following the next was assigned a lag of 2, and so
on. We tagged offscreen events using half steps. For example, an offscreen
event that occurred after the prior segment but before the just-watched segment
would be assigned a lag of -0.5.

@@

What might be driving participants to retrodict further and more accurately
into the unobserved past, compared with their predictions of the unobserved
future? By inspecting the video content, we noticed that characters frequently
referenced both past events and (planned or predicted) future events in their
spoken conversations. We wondered whether the characters' references might show
temporal asymmetries that might explain participants' behaviors. Across all of
the characters' conversations, and across all of the video segments from our
main experiment, we manually identified a total of 82 references to past or
future events (i.e., that occurred onscreen or offscreen before or after the
events depicted in the current segment; Figs.~\ref{fig:result3}A, \references
A, \characterRefs). Characters in our main experiment's stimulus tended to
reference the past (52 references) more than the future (30 references),
consistent with previous work~\citep{DemiEtal18}. References to the past were
also skewed to more temporally distant events compared with references to the
future (Figs.~\ref{fig:result3}B,~\references B,~\characterRefs). These
asymmetries also held for characters in the replication experiment's stimulus
(Fig.~\ref{fig:meta-analysis}). These observations indicate that the characters
in the stimulus display a ``preference'' for the past (versus future) in their
conversations. Might this asymmetry be driving the asymmetries in participants'
retrodictions versus predictions?

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.8\textwidth]{results3}

  \caption{\textbf{Characters' references drive participants' retrodiction and
  prediction performance.} \textbf{A. Illustration of annotation approach.} We
  manually annotated references to events in past or future segments in
  characters' spoken conversations. We matched each such reference with its
  corresponding storyline event (and its corresponding segment number for
  onscreen events, or half-step segment number for offscreen events). We then
  tracked the hit rate separately for referenced versus unreferenced events in
  participants' uncued retrodictions and predictions. \textbf{B. Reference rate
  as a function of lag.} Across all possible just-watched segments (lag 0), the
  bar heights denote the average proportions of events referenced in other past
  (orange, negative lags) or future (blue, positive lags) segments in our main
  experiment's stimulus. \textbf{C. Difference in hit rates between all events
  and unreferenced events.} To highlight the effect of characters' references
  to past and future events on participants' retrodictions and predictions,
  here we display the difference in across-segment mean hit rates between all
  events and unreferenced events, as a function of temporal distance (lag) to
  the just-watched segment. \textbf{D. Hit rates for unreferenced events.} The
  average response hit rates for unreferenced events are displayed as a
  function of temporal distance to the just-watched segment. Error bars denote
  bootstrapped 95\% confidence intervals. Panels B--D: colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{E. Hit rates and counts of
  referenced and unreferenced events.} As a function of temporal distance to
  the just-watched segment, the sub-panels display the across-segment mean
  numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red) and
  unreferenced (gray) events that participants hit (darker shading) or missed
  (lighter shading) in their uncued retrodictions (top sub-panel) and uncued
  predictions (bottom sub-panel). Intuitively, the widths of the rectangles at
  each lag denote the total number of events at each possible lag. The darker
  shading denotes the proportions of events that participants retrodicted or
  predicted, and the lighter shading denotes the proportions of events that
  participants ``missed'' in their responses. For an analogous presentation of
  results from the replication experiment, see Figure~\characterRefs.}

@@

  \caption{\textbf{Referenced events are associated with higher hit rates, but
  referring events are not.} \textbf{A. Illustration of annotation approach.}
  We extended the annotation procedure depicted in Figure~\ref{fig:result3}A to
  also label which events in our main experiment's stimuli \textit{contained}
  references to events in other segments. \textbf{B. Referenced versus
  referring events.} During event $i$, when a character makes a reference to
  another event ($j$), we define $i$ as the \textit{referring} event and $j$ as
  the \textit{referenced} event. \textbf{C. Referring rate as a function of
  lag.} Across all possible just-watched segments (lag 0), the bar heights
  denote the across-segment mean proportions of events containing references to
  events in other past (orange, negative lags) or future (blue, positive lags)
  segments in our main experiment's stimuli. The bar colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{D. Hit rates and counts of
  referenced, referring, and other events.} As a function of temporal distance
  to the just-watched segment, the sub-panels display the numbers ($x$-axes)
  and hit rates ($y$-axes) of referenced (red), referring (green), and other
  (gray) events that participants hit (darker shading) or missed (lighter
  shading) in their uncued retrodictions (top sub-panel) and uncued predictions
  (bottom sub-panel). For a display of analogous results from our replication
  experiment see Figure~\referringReferenced.}

@@

The preceding analyses show that when characters reference past or future
events, those referenced events, and other events that are temporally adjacent
to the referenced events, are more likely to be retrodicted and predicted. In
other words, referring to a past or future event in conversation leads to a
``boost'' in that event's hit rate. We wondered whether this boost was
bi-directional. In particular: when a character refers (during a
\textit{referring event}) to another event (i.e., the \textit{referenced
event}), does this boost only the referenced event's hit rate, or does the
referring event also receive a boost? We labeled each event as a ``referring
event,'' a ``referenced event,'' or a ``other event'' (i.e., not referring or
referenced; Fig.~\ref{fig:result5}A, B). We limited our analysis to references
to onscreen (explicit) events. Consistent with our analysis of the proportions
of referenced events (Fig.~\ref{fig:result3}B), the proportions of
\textit{referring} events exhibited a \textit{forward} temporal asymmetry
(Fig.~\ref{fig:result5}C). Controlling for absolute lag, we found that
referring events were associated with lower hit rates than referenced events in
our main experiment (uncued retrodiction: OR = 0.03, $Z = -4.81$, $p < 0.001$,
CI: 0.01 to 0.11; uncued prediction: OR = 0.04, $Z = -5.84$, $p < 0.001$, CI:
0.01 to 0.12; Fig.~\ref{fig:result5}D) and had no reliable differences in hit
rates compared with other events (uncued retrodiction: OR = 0.37, $Z = -1.46$,
$p = 0.15$, CI: 0.10 to 1.41; uncued prediction: OR = 2.16, $Z = 1.68$, $p =
0.09$, CI: 0.88 to 5.30). We also observed this phenomenon in our replication
experiment (referenced events, uncued retrodiction: OR = XXX, $Z = XXX$, $p =
XXX$, CI: XXX to XXX; referenced events, uncued prediction: OR = XXX, $Z =
XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued retrodiction: OR = XXX,
$Z = XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued prediction: OR =
XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; Fig.~\referringReferenced). Taken
together, this indicates that only referenced events received a hit rate boost
(relative to other events), suggesting that the retrodictive and predictive
benefits of references are directed (i.e., asymmetric).

@@

An additional ambiguous case arose in one main experiment participant's
responses pertaining to segment 12, storyline 2, whereby the participant
correctly identified an onscreen event that had not been included in our
original annotations. To account for this participant's response, we
retroactively added that event to our annotations of that segment. We also
identified and counted unmatched events in participants' responses (i.e.,
events that did not match any annotated events). Cases where the two coders'
independent scoring disagreed were resolved through discussions between the two
coders.

@@

In principle, a given event could receive multiple labels. For example, during
event $A$, a character might speak about another event, $B$, during which a
reference to a third event ($C$) was made. In this scenario, event $B$ could be
both a ``referring event'' ($B \rightarrow C$) \textit{and} a referenced event
($A \rightarrow B$). In practice, however, this scenario was quite rare,
accounting for only one out of a total of 30 onscreen events in our main
experiment and XXX events in our replication experiment.
